---
title: "New Zealand's COVID-19 testing"
author: "*Minh Chau Van Nguyen*"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
    source_code: embed
runtime: shiny
---

<style>
body, h1, h2, h3, h4, h5, h6{
  font-family: Helvetica;
}
table.dataTable.hover tbody tr:hover, table.dataTable.display tbody tr:hover
{
background-color: rgba(223, 177, 214, 0.5) !important;
}
.shiny-input-container > label {
margin-bottom: -15px;
margin-top:5px;
}
.selectize-input .item {
color: #1a6ecc;
}
</style>


```{r setup, include=FALSE}
library(readxl)         # read data file
library(dplyr)          # data manipulation
library(highcharter)    # produce chart  
library(knitr)          # produce html table
library(formattable)
library(DT)     # produce html table

fake_dt <- read_excel("Analyst_Lab_Data 2.xls")     # fake lab data
true_dt <- read_excel("Population_Data 2.xls")      # real population data
names(true_dt)[2] <- "SA22018_V1_00"                # rename SA22018_V1_00_NAME for merging purpose

# combine fake data and real data together, joined by SA22018_V1_00 and Ethnicity
# data excluding those that replied "Unknown" for Ethnicity (n = 1656 observations)
full_dt <- merge(fake_dt, true_dt, by = c("SA22018_V1_00", "Ethnicity"))

island <- c("South Island", "North Island", "South Island", "North Island",
            "North Island", "South Island", "Chatham Island", "South Island",
            "South Island", "South Island", "North Island", "North Island",
            "South Island", "South Island", "North Island", "North Island",
            "North Island", "North Island", "South Island", "South Island",
            "South Island", "North Island", "North Island", "North Island",
            "North Island", "South Island", "North Island", "South Island",
            "North Island", "North Island", "North Island", "South Island",
            "North Island", "North Island", "North Island", "North Island",
            "North Island", "South Island", "North Island", "North Island",
            "North Island", "South Island", "North Island", "North Island",
            "North Island", "South Island", "North Island", "North Island",
            "South Island", "North Island", "North Island", "North Island",
            "South Island", "North Island", "North Island", "South Island",
            "South Island", "North Island", "North Island", "South Island",
            "North Island", "North Island", "North Island", "South Island",
            "North Island", "North Island", "North Island")
```


Inputs {.sidebar}
-----------------------------------------------------------------------
```{r}
region_data <- data.frame(cbind(Region = levels(factor(full_dt$TA2018_V1_00_NAME)), 
                     Island = island))
selectInput(inputId = "Location", 
            label = h5("Terriotal Authority"), 
            choices = split(region_data$Region, region_data$Island),
            selected = "Napier City")
```

Key definitions:

<hr>

<div style = "font-size: 11.5px; margin-bottom:20px;">
$\star$ **Priotised Ethnicity** - "each person being allocated to a SINGLE ethnic group based on the ethnic groups they have identified with". The order is followed by M훮ori, Pacific, Asian and Other.

$\star$ **NZ Deprivation Index** - describes New Zealand's least to most deprived areas. A value of 10 on the index of deprivation indicates that the area is in the most deprived 10 percent of areas in New Zealand.

$\star$ **Population Estimate** - this is based on 2019 Census subnational population estimates.

</div>

Limitations:
<hr>

Row {data-height=750}
-----------------------------------------------------------------------

### **Chart** - showing the number of tests taken per day for each Ethnicity group at sub-Region/Area level within Region (Input). Hover over each bar for more information.
```{r}
# subset data for each Terriotal Authority/Region
# make this data reactive for use and reproducibility
reactive_data <- reactive({
  subset_data <- full_dt[full_dt$TA2018_V1_00_NAME == input$Location,]
  data <- subset_data[names(subset_data) %in% c("Ethnicity", "SA22018_V1_00_NAME", "Number_of_Tests", "SA2_average_NZDep2018", "Census_2019_Population_Estimate")]
  
  # Highchart tooltip doesn't like column names with special characters or spacing,
  # for example Number_of_Tests
  names(data)[c(2:5)] <- c("NZDep", "Area", "y", "Population")
  
  data$Ethnicity <- factor(data$Ethnicity)
  data$NZDep <- as.character(data$NZDep)
  data$Population <- as.character(round(data$Population))
  
  # rename Maori as M훮ori
  levels(data$Ethnicity) <- c("Asian", "M훮ori", "Other", "Pacific")
  data$Ethnicity <- factor(data$Ethnicity, levels = c("M훮ori", "Pacific", "Asian", "Other"))
  data
})

renderHighchart({
  data <- reactive_data()
  
  # Order data with respect to Ethnicity group otherwise highchart won't ouput the correct results
  data <- with(data, data[order(Ethnicity, Area, y, NZDep, Population),])
  
  data$Area <- factor(data$Area)
  
  hc <- highchart() %>%
    hc_chart(type = "column", inverted = TRUE, style = list(fontFamily = "Helvetica")) %>%
    hc_yAxis(title = list(text = "Number of tests per day", 
                          style = list(fontWeight = 'bold')),
             lineColor = "black", lineWidth = 1, reversedStacks = FALSE,
             labels = list(formatter = JS("function () {  return Highcharts.numberFormat(this.value, 0, '.', ',');}"))) %>%
    hc_xAxis(categories = levels(data$Ethnicity), 
           title = list(text = "Priotised Ethnicity group", 
                        style = list(fontWeight = 'bold')),
           lineColor = "black") %>%
  hc_plotOptions(column = list(
    stacking = "normal",
    borderColor = "#5ab1ef",
    borderWidth = 2,
    enableMouseTracking = TRUE))

  # Loop through each sub-Region/Area levels
  for(i in 1:length(levels(data$Area))){
    lvl <- levels(data$Area)[i]
    # Use list_parse to add external data to tooltip
    da <- list_parse(data %>% filter(Area==lvl) %>% select(y, pop = Population, name = NZDep))
    # show all series if data has less than or equal to 10 levels
    if(i <= 10){
      hc <- hc_add_series(hc, data = da, name = lvl, visible = TRUE)
    # show only 10 series if data has more than 10 levels
    }else{
      hc <- hc_add_series(hc, data = da, name = lvl, visible = FALSE)
    }
}

hc <- hc %>%
  hc_tooltip(formatter= JS("function () {
  return  '\u25CF ' + this.x +
  '<br /> <b>Area: </b>' + this.series.name  +
  '<br /> <b>Number of tests:</b> ' + Highcharts.numberFormat(this.point.y, -1, '.', ',') + ' tests' +
  '<br /> <b>Population estimate:</b> ' + this.point.pop + ' persons' +
  '<br /> <b>NZ Deprivation Index:</b> ' + this.point.name}")) %>%
  hc_legend(
    title = list(text = "Area <span style='font-size:9px; color: #666; font-weight: norma'>(Click to hide/show)</span>"),
    align = "left",
    verticalAlign = "top",
    layout = "vertical") 

hc
})
```


Row 
-----------------------------------------------------------------------

### **Table** - describes the total number of tests, total population estimates and average New Zealand Deprivation Index for each selected region (Input), grouped by Ethnicity.
```{r}
data_for_table <- reactive({
  data <- reactive_data()
  
  # calculate sum of Number of tests and sum of Population estimates, grouped by Ethnicity
  dt1 <- aggregate(cbind(Tests = data$y, Population = as.numeric(data$Population)), 
          by=list(Ethnicity = data$Ethnicity), FUN = sum)
  # calculate average of NZDep, grouped by Ethnicity
  dt2 <- aggregate(cbind(NZDep = as.numeric(data$NZDep)),
                   by = list(Ethnicity = data$Ethnicity), FUN = mean)
  
  dt2$NZDep <- round(dt2$NZDep)
  table_data <- data.frame(cbind(Location = rep(input$Location,4),
                      dt1, dt2[-1]))
  table_data
})


custom_color_tile <- function (...){
  data <- data_for_table()
    formatter("span",
              style = function(x) 
                style(display = "block", 
                      padding = "0 4px", 
                      `color` = "black", 
                      `border-radius` = "4px", 
                      `background-color` = ifelse(data$Population > data$Tests, "#B1CBEB", "none")))
}

DT::renderDataTable({
  m <- data_for_table() %>% 
    formattable(list(area(col=1:5) ~ custom_color_tile()))
   return(as.datatable(m, rownames = FALSE, 
                      colnames = c('Region', 'Priotised Ethnicity*', 'Number of tests', 'Population estimates (persons)*', 'NZ Deprivation Index*'),
                       caption = tags$caption(
        style = "caption-side:bottom; text-align:left; margin:-2px 0 0 5px; font-size:11.5px; font-style:italic;",
        tags$span("Rows highlighted in blue indicating the number of tests taken is smaller than the population estimate for the corresponding Ethnicity group and Location.", style = "color:blue;"), tags$br(), tags$span("*These measures were rounded to whole number.", style = "color:black;")),
        options = list(searching = FALSE, paging = FALSE, lengthChange = FALSE,  
                       bInfo = FALSE, columnDefs = list(list(className = 'dt-left', targets = "_all")))
          ))
})

```



