---
title: 'New Zealand''s fictitious COVID-19 flexdashboard'
runtime: shiny
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    # navbar:
    #   - { icon: "fa-github", 
    #       title: github, 
    #       href: "https://github.com/MinhChauVanNguyen/NZ_COVID-19", 
    #       align: right}
---

<style>
.navbar-brand {
margin-left:220px !important;
font-weight:bold;
}

.highcharts-series-group .highcharts-point {
  opacity: 0.6;
}

body, label, input, button, select, box, .h1, .h2, .h3, .h4, .h5, h1, h2, h3, h4, h5{
  font-family: Helvetica;
}

.caption1 {
font-size:90%;
}
table.dataTable  {
  font-size: 90%;
}

.shiny-input-container > label {
  margin-bottom: -15px;
}
</style>

<script>
(function (H) {
    H.wrap(H.Legend.prototype, 'colorizeItem', function (proceed, item, visible) {
        proceed.apply(this, Array.prototype.slice.call(arguments, 1));
        if(item.legendLine) {
            item.legendLine.attr({ stroke: item.color });
        }
        if(item.legendSymbol) {
            if ((item.options && item.options.marker) && item.legendSymbol.isMarker)
                item.legendSymbol.attr(item.pointAttribs());
            else
                item.legendSymbol.attr({ fill: item.color });
        }
    });
}(Highcharts));
</script>

    
```{r setup, include=FALSE}
library(knitr)
library(dplyr)          # data manipulation
library(highcharter)    # produce chart  
library(formattable)    # produce html table
library(DT)             # produce html table
library(shinyBS)        # info message 
library(shinyWidgets)

source("hc_theme_dark.R")
source("data_cleaning.R")
```



Inputs {.sidebar}
-----------------------------------------------------------------------

<div style = "margin-top:-10px;">
```{r}
selectInput(inputId = "Location", 
            label = "Territorial Authority", 
            choices = levels(full_data$TA2018_Name),
            selected = "Whangarei District")

#current_theme <- reactiveVal(TRUE)
renderUI({
  materialSwitch(inputId = "style", label = "Dark mode activated", value = TRUE)
})

observeEvent(input$style,{
  if(input$style == TRUE) {
    txt <- "Dark mode activated"
  } else {
    txt <- "Light mode activated"
  }
  updateCheckboxInput(session, "style", label = txt)
})

renderUI({
  # if (!is.null(input$style)) {
  #     current_theme(input$style)
  if(!is.null(input$style) & isTRUE(input$style)){
    includeCSS("www/theme-dark.css")
  }else{
    includeCSS("www/theme-light.css")
  }
   # }
})
```
</div>

<div style = "margin-top:10px;"></div>
Abstract

<hr>

<div style = "font-size: 85%; margin-bottom:20px;">

The dashboard displays graphs of COVID-19 (number of new positive confirmed cases within 6 months) data recorded at Statistical Area 2 (SA22018) level of geography by [Prioritised Ethnicity](https://www.moh.govt.nz/notebook/nbbooks.nsf/0/4b1c9119d4e5ea78cc257488000d9324/$FILE/presenting-ethnicity-2008.pdf). In addition to the [made up](https://github.com/MinhChauVanNguyen/NZ_COVID-19/blob/master/data_cleaning.R) number of people by Ethnicity and number of cases by Ethnicity features, the data has features that are based on real information, such as [Population estimates by SA2](https://www.stats.govt.nz/information-releases/statistical-area-1-dataset-for-2018-census-updated-march-2020) and [New Zealand Deprivation Index 2018](https://www.otago.ac.nz/wellington/departments/publichealth/research/hirp/otago020194.html). 

</div>



```{r}
actionLink(inputId = "info", 
           label = "Click for instructions on how to read the bubble chart on the right",
           style = 'font-size:85%;font-weight:bold; text-decoration:underline;',
           icon = icon("info"))

bsAlert(anchorId = "Info")

observeEvent(input$info, {
  createAlert(session, anchorId = "Info", title = "Please read", 
              content = HTML("<div class = alert-info role = alert style = 'text-align:left;'>
               <b>1. This chart plots Number of cases against total population by Ethnicity at SA2 levels <br>
               1. Labels of the bubbles are described by Statistic Area 2 (SA2) codes <br>
               2. Sizes of the bubbles correspond to NZ Deprivation Index values, where the bigger the bubble the higher the index.
                </b></div>"), append = FALSE, style = "info")
  #jqui_draggable(selector = '.alert-info')
})
```


<div style = "font-size: 85%; margin-top:20px;"></div>

Key definitions:

<hr>
<div style = "font-size: 85%; margin-bottom:20px;">
$\star$ **Prioritised Ethnicity** - *"each person being allocated to a SINGLE ethnic group based on the ethnic groups they have identified with"*. The order is followed by M훮ori, Pacific, Asian and Other.

$\star$ **NZ Deprivation Index** - describes New Zealand's least to most deprived areas. A value of 10 on the scale of deprivation indicates that the area is in the most deprived areas and vice versa. It is important to note that the data does contain some missing information regarding the Deprivation Index (n = 124 observations) which are shown as *NA* in the stacked bar graph and table, and excluded from the bubble chart. 
</div>


<footer>
<span style="font-size:12px;">*Start date: <span style="font-size:12px;">13$^{th}$ October, 2020*</span>
</footer>


Row 
-----------------------------------------------------------------------


### **Stacked Bar Chart** {.dark-box}

<div style="font-size:85%;">This bar chart plots the number of positive confirmed cases within 6 months for each Ethnicity group at SA2 level. Hover over each bar for more information.</div>


```{r}
# subset data for each Terriotal Authority/Region
# make this data reactive for use and reproducibility
reactive_data <- reactive({
  subset_data <- full_data[full_data$TA2018_Name == input$Location,]
  
  data <- subset_data[names(subset_data) %in% c("SA22018_Code", "SA22018_Name", "SA22018_NZDep", "Ethnicity", "Population_by_Ethnicity", "Number_of_cases")]

    # Highchart tooltip doesn't like column names with special characters or spacing, for example Number_of_Tests
  names(data) <- c("name", "Area", "NZDep", "Ethnicity", "Population", "y")

  data
})

renderHighchart({
  data <- reactive_data()
  
  data$NZDep <- as.character(data$NZDep)
  data$Population <- as.character(data$Population)
  
  # Order data with respect to Ethnicity group otherwise highchart won't output the correct results
  data <- with(data, data[order(Ethnicity, Area, y, NZDep, Population),])
  
  data$Area <- factor(data$Area)
  
  hc <- highchart() %>%
     hc_yAxis(title = list(text = "Number of confirmed positive cases", 
                          style = list(fontWeight = 'bold')),
             lineWidth = 1, reversedStacks = FALSE, tickColor = "white",
             labels = list(formatter = JS("function () {  return Highcharts.numberFormat(this.value, 0, '.', ',');}"))) %>%
    hc_chart(type = "column", height = 350, inverted = TRUE, 
             style = list(fontFamily = "Helvetica")) %>% 
    hc_xAxis(categories = levels(data$Ethnicity), 
           title = list(text = "Prioritised Ethnicity group", 
                        style = list(fontWeight = 'bold')),
           lineWidth = 1) %>%
    hc_plotOptions(
      column = list(
        stacking = "normal",
        borderColor = "#5ab1ef",
        borderWidth = 2,
        enableMouseTracking = TRUE)) 
  
  # Loop through each sub-Region/Area levels
  for(i in 1:length(levels(data$Area))){
    lvl <- levels(data$Area)[i]
    # Use list_parse to add external data to tooltip
    da <- list_parse(data %>% filter(Area==lvl) %>% select(y, pop = Population, name = NZDep))
    # show all series if data has less than or equal to 10 levels
    if(i <= 10){
      hc <- hc_add_series(hc, data = da, name = lvl, visible = TRUE)
    # show only 10 series if data has more than 10 levels
    }else{
      hc <- hc_add_series(hc, data = da, name = lvl, visible = FALSE)
    }
  }

  cols <-  c('#2f7ed8', '#ff99cc', '#00cc99', '#1aadce',
             '#ff9966', '#ffff00', '#77a1e5', '#9999ff', '#0099CC',
             '#CC66FF', '#ff6699', '#ffccff', '#0099FF', '#FF0099')
  
hc <- hc %>%
  hc_colors(cols) %>%
  hc_tooltip(
    useHTML = TRUE,
    formatter= JS("function () {
    var color = {M훮ori:'#5ac8c8', Pacific:'#ff83c1', Asian:'#7cb5ec', Other:'#FF6699'}
    return  '<span style =\"color:' + color[this.x] + ';\">' +  '\u25CF ' +  '<b>' + this.x + '</b>'  + '</span>' + 
    '<br /> <b>Area: </b>' + this.series.name  +
    '<br /> <b>Number of cases:</b> ' + Highcharts.numberFormat(this.point.y, -1, '.', ',') + ' cases' +
    '<br /> <b>Population by ethnicity:</b> ' + Highcharts.numberFormat(this.point.pop, -1, '.', ',') + ' persons' +
    '<br /> <b>NZ Deprivation Index:</b> ' + this.point.name }")) %>%
  hc_legend(
    align = "left",
    verticalAlign = "top",
    layout = "vertical"
  )

 if(isTRUE(input$style)){
   hc <- hc %>% hc_add_theme(hc_dark_theme()) %>%
     hc_legend(
       navigation = list(
         activeColor = "#00CCFF",
         style = list(
           color = "#FFCCFF"
      )
    ),
    title = list(text = "<span style='color:white;'>Area </span> <span style='font-size:9px;color:#FFCCFF; font-weight:bold;'>(Click on dots to hide/show)</span>"))%>%
  hc_tooltip(
    backgroundColor = '#34495e', borderWidth = 2,
    style = list(
        color = "white"
      )
  )
 } else {
    hc <- hc %>%
      hc_legend(title = list(
        text = "Area <span style='font-size:9px;color:#666; font-weight:bold;'>(Click on dots to hide/show)</span>")) %>%
      hc_add_theme(hc_light_theme())
 }

hc %>% hc_xAxis(tickColor = "none")
})
```


Row {data-height=650}
-----------------------------------------------------------------------

### **Bubble Chart** {.dark-box}

```{r}
renderHighchart({
 data <- reactive_data()
 
 names(data) <- c("name", "Area", "z", "Ethnicity", "x", "y")
 
 data <- data[!(data$z == "NA"),]
 
 data$z <- as.numeric(data$z)
 
hc <- highchart() %>%
  hc_chart(type = "bubble", plotBorderWidth = 1, 
           style = list(fontFamily = "Helvetica"))

if(isTRUE(input$style)){
  hc <- hc %>% hc_add_series(name = '<span style = color:white;">Ethnicity</span><br/><span style="font-size:9px; color:pink; font-weight:bold;">(Click on dots to hide/show)</span>', 
                color = 'none',
                marker = list(enabled = FALSE),
                events = list(legendItemClick = JS("function() {return false;}")))
  } else {
    hc <- hc %>%  hc_add_series(name = 'Ethnicity<br/><span style="font-size:9px; color:#666; font-weight:bold;">(Click on dots to hide/show)</span>', color = 'none',
                 marker = list(enabled = FALSE),
                 events = list(legendItemClick = JS("function() {return false;}")))
}

hc <- hc %>%
  hc_add_series(data = list_parse(data[data$Ethnicity == "M훮ori", ]), color = "#5ac8c8", name = "M훮ori") %>%
  hc_add_series(data = list_parse(data[data$Ethnicity == "Pacific", ]), color = "#ff83c1", name = "Pacific", visible = FALSE) %>%
  hc_add_series(data = list_parse(data[data$Ethnicity == "Asian", ]), name = "Asian", visible = FALSE) %>%
  hc_add_series(data = list_parse(data[data$Ethnicity == "Other", ]), color = "#FF6699", name = "Other", visible = FALSE) %>%
  hc_xAxis(title = list(text = 'Population estimate (persons)', 
                        style = list(fontWeight = 'bold')),
           labels = list(style = list(fontSize = '9px'),
                         formatter = JS("function () {  return Highcharts.numberFormat(this.value, 0, '.', ',');}")),
           gridLineWith = 1) %>%
  hc_yAxis(startOnTick = FALSE, endOnTick = FALSE,  # doesn't work in browser
           title = list(text = 'Number of positive confirmed cases', 
                       style = list(fontWeight = 'bold')),           
           labels = list(style = list(fontSize = '9px')),
           maxPadding = 0.2, gridLineWidth = 1) %>%
  hc_plotOptions(
    series = list(
      dataLabels = list(
        enabled = TRUE, format = '{point.name}',
        style = list(
          fontSize = '9px'
        )))) %>%
   hc_tooltip(useHTML = TRUE, formatter= JS("function () {
      return '<span style =\"color:' + this.point.series.color + ';font-weight:bold;\">\u25CF ' + this.point.Area + '</span>' + 
  '<br /> <b>Number of cases:</b> ' + Highcharts.numberFormat(this.point.y, -1, '.', ',') + ' cases' +
  '<br /> <b>Population by ethnicity:</b> ' + Highcharts.numberFormat(this.point.x, -1, '.', ',') + ' persons' +
  '<br /> <b>NZ Dep:</b> ' + this.point.z}")) %>%
  hc_legend(
     layout = 'horizontal',
     align =  'center',
     verticalAlign = 'top',
     floating = TRUE)

if(isTRUE(input$style)){
  hc <- hc %>% hc_add_theme(hc_dark_theme()) %>%
    hc_tooltip(
      backgroundColor = '#34495e', borderWidth = 2,
      style = list(
        color = "white"
      )
  ) 
} else {
  hc <- hc 
}

hc 

})
```


### **Table** {.dark-box}

```{r}
table_data <- reactive({
  subset_data <- full_data[full_data$TA2018_Name == input$Location,]
  
  # filter data
  subset_data <- subset_data %>% 
    mutate(Incidence = (Number_of_cases/(Population_by_Ethnicity*6)))
  
  subset_data <- subset_data[which(subset_data$Incidence > 0.005),]
  
  subset_data <- subset_data[,!(names(subset_data) %in% c("SA22018_Code", "Region", "SA22018_Total_Pop", "TA2018_Name"))]
    
  subset_data
})

custom_highlight <- function (dark_theme = TRUE){
  data <- table_data()
  formatter("span",
            style = function(x) 
              style(display = "block",
                    padding = "0 4 px",
                    `border-radius` = "4px",
                    `color` = ifelse(dark_theme == TRUE, "white", "black"),
                    `background-color` = 
                      ifelse(dark_theme == TRUE  & data$Incidence > 0.05, "#2f7ed8", ifelse(dark_theme == FALSE & data$Incidence > 0.05, "#B1CBEB", "none"))))
  }
     

DT::renderDataTable({
  
  validate(
     need(nrow(table_data())>0, "Sorry, there is no data availale based on the indicated filter for the selected district/city.")
   )
  
  options <- list(
       searching = FALSE, paging = FALSE, lengthChange = FALSE, bInfo = FALSE,
       columnDefs = list(list(className = 'dt-left', targets = "_all")))
  
  data <- table_data() %>% 
    formattable(list(
      area(col = 1:6) ~ custom_highlight(ifelse(isTRUE(input$style), TRUE, FALSE))))
     data$Incidence <- percent(data$Incidence, 1)
  
     my_table <- return(as.datatable(
        data,
        fillContainer = TRUE, rownames = FALSE,
        colnames = c('Area', 'NZDep', 'Ethnicity', 'Population', 'Cases', 'Incidence rate*'),
        options = options,
        caption = tags$caption(
            tags$span("Table summarising filtered data where the incidence rate are larger than 0.5% (or 5 cases per 100 persons-month) for", 
                      class = "caption1"),
            tags$span(paste(input$Location, "."), class = "location"),
            tags$br(),
            tags$span("*Rows highlighted in blue indicating the rates are extreme - more than 5 cases per 100 persons-month.", class = "caption2")
        )) %>% formatStyle('SA22018_Name', target = 'row', 
                  backgroundColor = ifelse(isTRUE(input$style), "#34495e", "none"))
      )

  my_table
})
```
