---
title: "*New Zealand's COVID-19 testing dashboard for Ministry of Health*"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
runtime: shiny
---

<style>
.navbar-brand {
margin-left:220px !important;
}

.shiny-output-error-validation {
    color: #ff0000;
    font-weight: bold;
    margin-top:10px;
    text-align:center;
}

body, h1, h2, h3, h4, h5, h6{
  font-family: Helvetica;
}

table.dataTable.hover tbody tr:hover, table.dataTable.display tbody tr:hover{
  background-color: rgba(223, 177, 214, 0.5) !important;
}

table.dataTable  {
  font-size: 90%;
}

table.dataTable.no-footer {
  border-bottom: 1px solid #ddd;
}

.shiny-input-container > label {
  margin-bottom: -15px;
  margin-top:5px;
}

.selectize-input .item {
color: #1a6ecc;
}

.alert-info {
  background-color:#042f66;
  color:white;
  font-size:bold;
  font-size:90%;
}
</style>


```{r setup, include=FALSE}
library(readxl)         # read data file
library(dplyr)          # data manipulation
library(highcharter)    # produce chart  
library(formattable)    # produce html table
library(DT)             # produce html table
library(shinyBS)        # info message 
#library(shinyjqui)

fake_dt <- read_excel("Analyst_Lab_Data 2.xls")     # fake lab data (n = 9758)
true_dt <- read_excel("Population_Data 2.xls")      # real population data (n = 8608)
names(true_dt)[2] <- "SA22018_V1_00"                # rename SA22018_V1_00_NAME for merging purpose
# combine fake data and real data together, joined by SA22018_V1_00 and Ethnicity
# data excluding those that replied "Unknown" for Ethnicity (n = 1656 observations)
# nrow(fake_dt[fake_dt$Ethnicity == "Unknown",])
# n = 8102
full_dt <- merge(fake_dt, true_dt, by = c("SA22018_V1_00", "Ethnicity"))

full_dt$Census_2019_Population_Estimate <- round(full_dt$Census_2019_Population_Estimate)

full_dt$Ethnicity <- factor(full_dt$Ethnicity)

# rename Maori as Māori
levels(full_dt$Ethnicity) <- c("Asian", "Māori", "Other", "Pacific")
full_dt$Ethnicity <- factor(full_dt$Ethnicity, levels = c("Māori", "Pacific", "Asian", "Other"))

# map each region to South and North Islands
island <- c("South Island", "North Island", "South Island", "North Island",
            "North Island", "South Island", "Chatham Island", "South Island",
            "South Island", "South Island", "North Island", "North Island",
            "South Island", "South Island", "North Island", "North Island",
            "North Island", "North Island", "South Island", "South Island",
            "South Island", "North Island", "North Island", "North Island",
            "North Island", "South Island", "North Island", "South Island",
            "North Island", "North Island", "North Island", "South Island",
            "North Island", "North Island", "North Island", "North Island",
            "North Island", "South Island", "North Island", "North Island",
            "North Island", "South Island", "North Island", "North Island",
            "North Island", "South Island", "North Island", "North Island",
            "South Island", "North Island", "North Island", "North Island",
            "South Island", "North Island", "North Island", "South Island",
            "South Island", "North Island", "North Island", "South Island",
            "North Island", "North Island", "North Island", "South Island",
            "North Island", "North Island", "North Island")
```


Inputs {.sidebar}
-----------------------------------------------------------------------
```{r}
full_dt$TA2018_V1_00_NAME <- factor(full_dt$TA2018_V1_00_NAME)

region_data <- data.frame(cbind(Region = levels(full_dt$TA2018_V1_00_NAME), 
                     Island = island))

selectInput(inputId = "Location", 
            label = h5("Territorial Authority"), 
            choices = split(region_data$Region, region_data$Island),
            selected = "Whangarei District")
```

Key definitions:

<hr>

<div style = "font-size: 85%; margin-bottom:30px;">
$\star$ **Prioritised Ethnicity** - *"each person being allocated to a SINGLE ethnic group based on the ethnic groups they have identified with"*. The order is followed by Māori, Pacific, Asian and Other.

$\star$ **NZ Deprivation Index** - describes New Zealand's least to most deprived areas. A value of 10 on the scale of deprivation indicates that the area is in the most deprived areas in New Zealand and vice versa.

$\star$ **Population Estimate** - this is based on 2019 Census population estimates recorded at Statistical Area 2 (SA2) geographical levels.

</div>

```{r}
actionLink(inputId = "info", "Click for instructions on how to read the bubble chart on the right", style='font-size:85%;font-weight:bold; text-decoration:underline;', icon = icon("info"))

bsAlert(anchorId = "Info")

observeEvent(input$info, {
  createAlert(session, anchorId = "Info", title = "Please read", 
              content = HTML("<div class = alert-info role = alert style = 'text-align:left;'>
               <b>1. This chart plots Number of tests against population estimates at SA2 levels <br>
               1. Labels of the bubbles are described by Statistic Area 2 (SA2) codes <br>
               2. Sizes of the bubbles correspond to NZ Deprivation Index values, where the bigger the bubble the higher the index.
                </b></div>"), append = FALSE, style = "info")
  #jqui_draggable(selector = '.alert-info')
})
```

<div style = "margin-top:30px";">Limitations of data analysis:</div>
<hr>

<div style = "font-size: 85%;">

$\star$ The data excludes those that are categorised as "Unknown" for Ethnicity (n = 1656 observations) hence doesn't take these individuals into account.

$\star$ The period of time for when the cumulative number of tests are assumed to be per month and taken during the previous month (September, 2020). 

</div>


<footer>
<span style="font-size:12px;">*Start date: <span style="font-size:12px;">13$^{th}$ October, 2020*</span>
</footer>

Row 
-----------------------------------------------------------------------


### **Stacked Bar Chart** 

<div style="font-size:85%;">This bar chart displays the number of tests taken per day for each Ethnicity group at sub-Region/Area level within Districts/Cities (Input). Hover over each bar for more information.</div>

```{r}
# subset data for each Terriotal Authority/Region
# make this data reactive for use and reproducibility
reactive_data <- reactive({
  subset_data <- full_dt[full_dt$TA2018_V1_00_NAME == input$Location,]
  data <- subset_data[names(subset_data) %in% c("Ethnicity", "SA22018_V1_00", "SA22018_V1_00_NAME", "Number_of_Tests", "SA2_average_NZDep2018", "Census_2019_Population_Estimate")]

    # Highchart tooltip doesn't like column names with special characters or spacing, for example Number_of_Tests
  names(data) <- c("name", "Ethnicity", "NZDep", "Area", "y", "Population")

  data
})

renderHighchart({
  data <- reactive_data()
  
  data$NZDep <- as.character(data$NZDep)
  data$Population <- as.character(data$Population)
  
  # Order data with respect to Ethnicity group otherwise highchart won't ouput the correct results
  data <- with(data, data[order(Ethnicity, Area, y, NZDep, Population),])
  
  data$Area <- factor(data$Area)
  
  hc <- highchart() %>%
    hc_chart(type = "column", height = 350, inverted = TRUE, style = list(fontFamily = "Helvetica")) %>%
    hc_yAxis(title = list(text = "Number of tests", 
                          style = list(fontWeight = 'bold')),
             lineColor = "black", lineWidth = 1, reversedStacks = FALSE,
             labels = list(formatter = JS("function () {  return Highcharts.numberFormat(this.value, 0, '.', ',');}"))) %>%
    hc_xAxis(categories = levels(data$Ethnicity), 
           title = list(text = "Prioritised Ethnicity group", 
                        style = list(fontWeight = 'bold')),
           lineColor = "black") %>%
  hc_plotOptions(column = list(
    stacking = "normal",
    borderColor = "#5ab1ef",
    borderWidth = 2,
    enableMouseTracking = TRUE)) 
  
  # Loop through each sub-Region/Area levels
  for(i in 1:length(levels(data$Area))){
    lvl <- levels(data$Area)[i]
    # Use list_parse to add external data to tooltip
    da <- list_parse(data %>% filter(Area==lvl) %>% select(y, pop = Population, name = NZDep))
    # show all series if data has less than or equal to 10 levels
    if(i <= 10){
      hc <- hc_add_series(hc, data = da, name = lvl, visible = TRUE)
    # show only 10 series if data has more than 10 levels
    }else{
      hc <- hc_add_series(hc, data = da, name = lvl, visible = FALSE)
    }
  }
  
hc <- hc %>%
  hc_tooltip(
    useHTML = TRUE,
    formatter= JS("function () {
    var color = {Māori:'#5ac8c8', Pacific:'#ff83c1', Asian:'#8c63aa', Other:'#7cb5ec'}
    return  '<span style =\"color:' + color[this.x] + ';\">' +  '\u25CF ' +  '<b>' + this.x + '</b>'  + '</span>' + 
    '<br /> <b>Area: </b>' + this.series.name  +
    '<br /> <b>Number of tests:</b> ' + Highcharts.numberFormat(this.point.y, -1, '.', ',') + ' tests' +
    '<br /> <b>Population estimate:</b> ' + Highcharts.numberFormat(this.point.pop, -1, '.', ',') + ' persons' +
    '<br /> <b>NZ Deprivation Index:</b> ' + this.point.name }")) %>%
  hc_legend(
    title = list(text = "Area <span style='font-size:9px; color: #666; font-weight:bold;'>(Click on dots to hide/show)</span>"),
    align = "left",
    verticalAlign = "top",
    layout = "vertical") 

hc
})
```



Row {data-height=650}
-----------------------------------------------------------------------

### **Bubble Chart** 
```{r}
renderHighchart({
 data <- reactive_data()
 names(data) <- c("name", "Ethnicity", "z", "Area", "y", "x")
 
 highchart() %>%
  hc_chart(type = "bubble", plotBorderWidth = 1) %>%
  hc_add_series(name = 'Ethnicity<br/><span style="font-size:9px; color:#666; font-weight:bold;">(Click on dots to hide/show)</span>', color = 'none',
                marker = list(enabled = FALSE),
                events = list(legendItemClick = JS("function() {return false;}"))) %>%
  hc_add_series(data = list_parse(data[data$Ethnicity == "Māori", ]), color = "#5ac8c8", name = "Māori") %>%
  hc_add_series(data = list_parse(data[data$Ethnicity == "Pacific", ]), color = "#ff83c1", name = "Pacific", visible = FALSE) %>%
  hc_add_series(data = list_parse(data[data$Ethnicity == "Asian", ]), color = "#8c63aa", name = "Asian", visible = FALSE) %>%
  hc_add_series(data = list_parse(data[data$Ethnicity == "Other", ]), name = "Other", visible = FALSE) %>%
  hc_xAxis(title = list(text = 'Population estimate (persons)', 
                        style = list(fontWeight = 'bold')),
           labels = list(style = list(fontSize = '9px'),
                         formatter = JS("function () {  return Highcharts.numberFormat(this.value, 0, '.', ',');}")),
           gridLineWith = 1) %>%
  hc_yAxis(startOnTick = FALSE, endOnTick = FALSE,  # doesn't work in browser
           title = list(text = 'Number of tests', 
                       style = list(fontWeight = 'bold')),           
           labels = list(style = list(fontSize = '9px')),
           maxPadding = 0.2, gridLineWidth = 1) %>%
  hc_plotOptions(
    #lang = list(thousandsSep = ','),
    series = list(
      dataLabels = list(
        enabled = TRUE, format = '{point.name}',
        style = list(
          fontSize = '9px'
        )))) %>%
   hc_tooltip(useHTML = TRUE,
   formatter= JS("function () {
  return '<span style =\"color:' + this.point.series.color + ';font-weight:bold;\">\u25CF ' + this.point.Area + '</span>' + 
  '<br /> <b>Number of tests:</b> ' + Highcharts.numberFormat(this.point.y, -1, '.', ',') + ' tests' +
  '<br /> <b>Population estimate:</b> ' + Highcharts.numberFormat(this.point.x, -1, '.', ',') + ' persons' +
  '<br /> <b>NZ Dep:</b> ' + this.point.z}")) %>%
  hc_legend(
     layout = 'horizontal',
     align =  'center',
     verticalAlign ='top',
     floating = TRUE) 
   #hc_size(height = 250)
})
```


### **Table**

```{r}
table_data <- reactive({
  data <- full_dt[full_dt$TA2018_V1_00_NAME == input$Location,]
  
  # filter data
  subset_data <- data[which(data$Number_of_Tests < data$Census_2019_Population_Estimate & data$SA2_average_NZDep2018 > 5),]
  
  subset_data <- subset_data[,!(names(subset_data) %in% c("SA22018_V1_00", "SA2_average_NZDep2018_score"))]
  
  subset_data <- subset_data %>% 
    mutate(Rate = Number_of_Tests/Census_2019_Population_Estimate) %>%
    select(SA22018_V1_00_NAME, Ethnicity, SA2_average_NZDep2018, Number_of_Tests, Census_2019_Population_Estimate, Rate)
  
  subset_data
})

custom_color_tile <- function (...){
  data <- table_data()
    formatter("span",
              style = function(x) 
                style(display = "block", 
                      padding = "0 4px", 
                      `color` = "black", 
                      `border-radius` = "4px", 
                      `background-color` = ifelse(data$Rate < 0.2, "#B1CBEB", "none")))
}

# bottom caption
js <- c(
  "function(settings){",
  "  var datatable = settings.oInstance.api();",
  "  var table = datatable.table().node();",
  "  var caption = 'Rows highlighted in blue indicating the test rates are less than 20 tests per 100 persons.'",
  "  $(table).append('<caption style=\"caption-side:bottom; text-align:left; margin:-2px 0 0 5px; font-size:11.5px; font-style:italic;color:#1a6ecc;\">' + caption + '</caption>');",
  "}"
)

DT::renderDataTable({
  validate(
     need(nrow(table_data())>0, "Sorry, there is no data availale based on the indicated filter for this region.")
   )
  data <- table_data() %>% formattable(
    list(area(col=1:6) ~ custom_color_tile()))

  data$Rate <- percent(data$Rate, 1)
  
   return(as.datatable(
     data, 
     fillContainer = TRUE,
     rownames = FALSE,
     colnames = c('Area', 'Ethnicity', 'NZDep', 'Total tests', 'Population estimate', 'Test rate per 100 persons'),
     caption = tags$caption(
        style = "color:black; margin:0 0 -15px 0; font-size:90%;",
        tags$span("Table summarising filtered testing data where the number of tests are smaller than the population estimates with a NZDep larger than 5 and grouped by ethnicity at SA2 levels, for"), tags$span(input$Location, style = 'color:#1a6ecc;font-weight:bold;')),
     options = list(
       drawCallback = JS(js),
       searching = FALSE, paging = FALSE, lengthChange = FALSE, bInfo = FALSE,
       columnDefs = list(list(className = 'dt-left', targets = "_all")))
    )
    )  
})
```

