---
title: 'New Zealand''s COVID-19 testing dashboard for Ministry of Health'
resource_files:
- 2018-SA1-dataset-dwellings-AucklandRegion_updated_11-6-20.xlsx
- 2018-SA1-dataset-dwellings-BayofPlentyRegion_updated_11-6-20.xlsx
- 2018-SA1-dataset-dwellings-CanterburyRegion_updated_11-6-20.xlsx
- 2018-SA1-dataset-dwellings-GisborneRegion_updated_11-6-20.xlsx
- 2018-SA1-dataset-dwellings-Hawke'sBayRegion_updated_11-6-20.xlsx
- 2018-SA1-dataset-dwellings-Manawatu-WanganuiRegion_updated_11-6-20.xlsx
- 2018-SA1-dataset-dwellings-MarlboroughRegion_updated_11-6-20.xlsx
- 2018-SA1-dataset-dwellings-NelsonRegion_updated_11-6-20.xlsx
- 2018-SA1-dataset-dwellings-OtagoRegion_updated_11-6-20.xlsx
- 2018-SA1-dataset-dwellings-NorthlandRegion_updated_11-6-20.xlsx
- 2018-SA1-dataset-dwellings-SouthlandRegion_updated_11-6-20.xlsx
- 2018-SA1-dataset-dwellings-TaranakiRegion_updated_11-6-20.xlsx
- 2018-SA1-dataset-dwellings-TasmanRegion_updated_11-6-20.xlsx
- 2018-SA1-dataset-dwellings-WaikatoRegion_updated_11-6-20.xlsx
- 2018-SA1-dataset-dwellings-WellingtonRegion_updated_11-6-20.xlsx
- 2018-SA1-dataset-dwellings-WestCoastRegion_updated_11-6-20.xlsx
runtime: shiny
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    #css: www/theme-dark.css
---

<style>
.navbar-brand {
margin-left:220px !important;
font-weight:bold;
}

.highcharts-series-group .highcharts-point {
  opacity: 0.6;
}

.shiny-output-error-validation {
    color: #ff0000;
    font-weight: bold;
    margin-top:10px;
    text-align:center;
}

body, label, input, button, select, box, .h1, .h2, .h3, .h4, .h5, h1, h2, h3, h4, h5{
  font-family: Helvetica;
}

table.dataTable  {
  font-size: 90%;
}

.shiny-input-container > label {
  margin-bottom: -15px;
  margin-top:5px;
}

.alert-info {
  background-color:#042f66;
  color:white;
  font-size:bold;
  font-size:90%;
}

.selectize-input .item {
color: #1a6ecc;
}
<!-- .first-box { -->
<!--   background-color: blue; -->
<!--   color: white; -->
<!-- } -->

</style>

<script>
(function (H) {
    H.wrap(H.Legend.prototype, 'colorizeItem', function (proceed, item, visible) {
        proceed.apply(this, Array.prototype.slice.call(arguments, 1));
        if(item.legendLine) {
            item.legendLine.attr({ stroke: item.color });
        }
        if(item.legendSymbol) {
            if ((item.options && item.options.marker) && item.legendSymbol.isMarker)
                item.legendSymbol.attr(item.pointAttribs());
            else
                item.legendSymbol.attr({ fill: item.color });
        }
    });
}(Highcharts));
</script>

```{r setup, include=FALSE}
library(dplyr)          # data manipulation
library(highcharter)    # produce chart  
library(formattable)    # produce html table
library(DT)             # produce html table
library(shinyBS)        # info message 

source("data_cleaning.R")
source("hc_theme_dark.R")
```


Inputs {.sidebar}
-----------------------------------------------------------------------
```{r}
selectInput(inputId = "Location", 
            label = h5("Territorial Authority"), 
            choices = levels(full_data$TA2018_Name),
            selected = "Whangarei District")
```

Abstract

<hr>

<div style = "font-size: 85%; margin-bottom:30px;">

The dashboard displays graphs of made-up COVID-19 (number of new positive confirmed cases within 6 months) data recorded at Statistical Area 2 (SA22018) level of geography by [Prioritised Ethnicity](https://www.moh.govt.nz/notebook/nbbooks.nsf/0/4b1c9119d4e5ea78cc257488000d9324/$FILE/presenting-ethnicity-2008.pdf). The data consists of other features that are based on real information, such as [Population estimates by SA2](https://www.stats.govt.nz/information-releases/statistical-area-1-dataset-for-2018-census-updated-march-2020) and [New Zealand Deprivation Index 2018](https://www.otago.ac.nz/wellington/departments/publichealth/research/hirp/otago020194.html). For information on the algorithms used for generating random numbers for the population by Ethnicity based on the given Total population by SA2, please refer to the [following](https://github.com/MinhChauVanNguyen/NZ_COVID-19/blob/master/data_cleaning.R).

</div>



```{r}
#current_theme <- reactiveVal(TRUE)

renderUI({
  checkboxInput(inputId = "style", label = "Dark mode activated", value = TRUE)
})

observeEvent(input$style,{
  if(input$style == TRUE) {
    txt <- "Dark mode activated"
  } else {
    txt <- "Light mode activated"
  }
  updateCheckboxInput(session, "style", label = txt)
})

renderUI({
  # if (!is.null(input$style)) {
  #     current_theme(input$style)
  if(!is.null(input$style) & isTRUE(input$style)){
    includeCSS("www/theme-dark.css")
  }else{
    includeCSS("www/theme-light.css")
  }
   # }
})

actionLink(inputId = "info", 
           label = "Click for instructions on how to read the bubble chart on the right",
           style = 'font-size:85%;font-weight:bold; text-decoration:underline;', 
           icon = icon("info"))

bsAlert(anchorId = "Info")

observeEvent(input$info, {
  createAlert(session, anchorId = "Info", title = "Please read", 
              content = HTML("<div class = alert-info role = alert style = 'text-align:left;'>
               <b>1. This chart plots Number of tests against population estimates at SA2 levels <br>
               1. Labels of the bubbles are described by Statistic Area 2 (SA2) codes <br>
               2. Sizes of the bubbles correspond to NZ Deprivation Index values, where the bigger the bubble the higher the index.
                </b></div>"), append = FALSE, style = "info")
  #jqui_draggable(selector = '.alert-info')
})
```


Key definitions:

<hr>
<div style = "font-size: 85%; margin-bottom:30px;">
$\star$ **Prioritised Ethnicity** - *"each person being allocated to a SINGLE ethnic group based on the ethnic groups they have identified with"*. The order is followed by M훮ori, Pacific, Asian and Other.

$\star$ **NZ Deprivation Index** - describes New Zealand's least to most deprived areas. A value of 10 on the scale of deprivation indicates that the area is in the most deprived areas in New Zealand and vice versa.
</div>


<footer>
<span style="font-size:12px;">*Start date: <span style="font-size:12px;">13$^{th}$ October, 2020*</span>
</footer>


Row 
-----------------------------------------------------------------------


### **Stacked Bar Chart** {.dark-box}

<div style="font-size:85%;">This bar chart displays the number of tests for each Ethnicity group at sub-Region/Area level within Districts/Cities (Input). Hover over each bar for more information.</div>


```{r}
# subset data for each Terriotal Authority/Region
# make this data reactive for use and reproducibility
reactive_data <- reactive({
  subset_data <- full_data[full_data$TA2018_Name == input$Location,]
  
  data <- subset_data[names(subset_data) %in% c("SA22018_Code", "SA22018_Name", "SA22018_NZDep", "Ethnicity", "Population_by_Ethnicity", "Number_of_cases")]

    # Highchart tooltip doesn't like column names with special characters or spacing, for example Number_of_Tests
  names(data) <- c("name", "Area", "NZDep", "Ethnicity", "Population", "y")

  data
})

renderHighchart({
  data <- reactive_data()
  
  data$NZDep <- as.character(data$NZDep)
  data$Population <- as.character(data$Population)
  
  # Order data with respect to Ethnicity group otherwise highchart won't output the correct results
  data <- with(data, data[order(Ethnicity, Area, y, NZDep, Population),])
  
  data$Area <- factor(data$Area)
  
  hc <- highchart() %>%
    hc_chart(type = "column", height = 350, inverted = TRUE, style = list(fontFamily = "Helvetica")) %>%
    hc_yAxis(title = list(text = "Number of confirmed positive cases", 
                          style = list(fontWeight = 'bold')),
             lineWidth = 1, reversedStacks = FALSE,
             labels = list(formatter = JS("function () {  return Highcharts.numberFormat(this.value, 0, '.', ',');}"))) %>%
    hc_xAxis(categories = levels(data$Ethnicity), 
           title = list(text = "Prioritised Ethnicity group", 
                        lineWidth = 1, lineColor = "black",
                        style = list(fontWeight = 'bold'))) %>%
    hc_plotOptions(
    column = list(
      stacking = "normal",
      borderColor = "#5ab1ef",
      borderWidth = 2,
      enableMouseTracking = TRUE)) 
  
  # Loop through each sub-Region/Area levels
  for(i in 1:length(levels(data$Area))){
    lvl <- levels(data$Area)[i]
    # Use list_parse to add external data to tooltip
    da <- list_parse(data %>% filter(Area==lvl) %>% select(y, pop = Population, name = NZDep))
    # show all series if data has less than or equal to 10 levels
    if(i <= 10){
      hc <- hc_add_series(hc, data = da, name = lvl, visible = TRUE)
    # show only 10 series if data has more than 10 levels
    }else{
      hc <- hc_add_series(hc, data = da, name = lvl, visible = FALSE)
    }
  }

  cols <-  c('#2f7ed8', '#ff99cc', '#8bbc21', '#00cc99', '#1aadce',
        '#492970', '#f28f43', '#77a1e5', '#c42525', '#a6c96a')
  
hc <- hc %>%
  hc_colors(cols) %>%
  hc_tooltip(
    useHTML = TRUE,
    formatter= JS("function () {
    var color = {M훮ori:'#5ac8c8', Pacific:'#ff83c1', Asian:'#7cb5ec', Other:'ff6699'}
    return  '<span style =\"color:' + color[this.x] + ';\">' +  '\u25CF ' +  '<b>' + this.x + '</b>'  + '</span>' + 
    '<br /> <b>Area: </b>' + this.series.name  +
    '<br /> <b>Number of cases:</b> ' + Highcharts.numberFormat(this.point.y, -1, '.', ',') + ' cases' +
    '<br /> <b>Population by ethnicity:</b> ' + Highcharts.numberFormat(this.point.pop, -1, '.', ',') + ' persons' +
    '<br /> <b>NZ Deprivation Index:</b> ' + this.point.name }")) %>%
  hc_legend(
    align = "left",
    verticalAlign = "top",
    layout = "vertical"
  )

  
 if(isTRUE(input$style)){
   hc <- hc %>% hc_add_theme(hc_dark_theme()) %>%
     hc_legend(
       navigation = list(
         activeColor = "#00CCFF",
         style = list(
           color = "#FFCCFF"
      )
    ),
    title = list(text = "<span style='color:white;'>Area </span> <span style='font-size:9px;color:#FFCCFF; font-weight:bold;'>(Click on dots to hide/show)</span>"))%>%
  hc_tooltip(
    backgroundColor = '#34495e', borderWidth = 2,
    style = list(
        color = "white"
      )
  )
 } else {
    hc <- hc %>%
      hc_legend(title = list(
        text = "Area <span style='font-size:9px;color:#666; font-weight:bold;'>(Click on dots to hide/show)</span>")) %>%
      hc_add_theme(hc_light_theme())
 }

hc
})
```


Row {data-height=650}
-----------------------------------------------------------------------

### **Bubble Chart** {.dark-box}
```{r}
renderHighchart({
 data <- reactive_data()
 
 names(data) <- c("name", "Area", "z", "Ethnicity", "x", "y")
 
hc <- highchart() %>%
  hc_chart(type = "bubble", plotBorderWidth = 1)

if(isTRUE(input$style)){
  hc <- hc %>% hc_add_series(name = '<span style = color:white;">Ethnicity</span><br/><span style="font-size:9px; color:pink; font-weight:bold;">(Click on dots to hide/show)</span>', 
                color = 'none',
                marker = list(enabled = FALSE),
                events = list(legendItemClick = JS("function() {return false;}")))
} else {
  hc <- hc %>%   hc_add_series(name = 'Ethnicity<br/><span style="font-size:9px; color:#666; font-weight:bold;">(Click on dots to hide/show)</span>', color = 'none',
                 marker = list(enabled = FALSE),
                 events = list(legendItemClick = JS("function() {return false;}")))
}

hc <- hc %>%
  hc_add_series(data = list_parse(data[data$Ethnicity == "M훮ori", ]), color = "#5ac8c8", name = "M훮ori") %>%
  hc_add_series(data = list_parse(data[data$Ethnicity == "Pacific", ]), color = "#ff83c1", name = "Pacific", visible = FALSE) %>%
  hc_add_series(data = list_parse(data[data$Ethnicity == "Asian", ]), name = "Asian", visible = FALSE) %>%
  hc_add_series(data = list_parse(data[data$Ethnicity == "Other", ]), color = "#FF6699", name = "Other", visible = FALSE) %>%
  hc_xAxis(title = list(text = 'Population estimate (persons)', 
                        style = list(fontWeight = 'bold')),
           labels = list(style = list(fontSize = '9px'),
                         formatter = JS("function () {  return Highcharts.numberFormat(this.value, 0, '.', ',');}")),
           gridLineWith = 1) %>%
  hc_yAxis(startOnTick = FALSE, endOnTick = FALSE,  # doesn't work in browser
           title = list(text = 'Number of positive confirmed cases', 
                       style = list(fontWeight = 'bold')),           
           labels = list(style = list(fontSize = '9px')),
           maxPadding = 0.2, gridLineWidth = 1) %>%
  hc_plotOptions(
    series = list(
      dataLabels = list(
        enabled = TRUE, format = '{point.name}',
        style = list(
          fontSize = '9px'
        )))) %>%
   hc_tooltip(useHTML = TRUE, formatter= JS("function () {
      return '<span style =\"color:' + this.point.series.color + ';font-weight:bold;\">\u25CF ' + this.point.Area + '</span>' + 
  '<br /> <b>Number of cases:</b> ' + Highcharts.numberFormat(this.point.y, -1, '.', ',') + ' cases' +
  '<br /> <b>Population by ethnicity:</b> ' + Highcharts.numberFormat(this.point.x, -1, '.', ',') + ' persons' +
  '<br /> <b>NZ Dep:</b> ' + this.point.z}")) %>%
  hc_legend(
     layout = 'horizontal',
     align =  'center',
     verticalAlign = 'top',
     floating = TRUE)

if(isTRUE(input$style)){
  hc <- hc %>% hc_add_theme(hc_dark_theme()) %>%
    hc_tooltip(
      backgroundColor = '#34495e', borderWidth = 2,
      style = list(
        color = "white"
      )
  )
} else {
  hc <- hc
}

hc
})
```


### **Table** {.dark-box}

```{r}
table_data <- reactive({
  subset_data <- full_data[full_data$TA2018_Name == input$Location,]
  
  # filter data
  subset_data <- subset_data %>% 
    mutate(Incidence = (Number_of_cases/(Population_by_Ethnicity*6)))
  
  subset_data <- subset_data[which(subset_data$Incidence > 0.005),]
  
    subset_data <- subset_data[,!(names(subset_data) %in% c("SA22018_Code", "Region", "SA22018_Total_Pop", "TA2018_Name"))]
    
  subset_data
})


custom_dark <- function (...){
    data <- table_data()
    formatter("span",
              style = function(x) 
                style(display = "block", 
                      padding = "0 4px", 
                      `color` = "white", 
                      `border-radius` = "4px", 
                      `background-color` = ifelse(data$Incidence > 0.05, "#2f7ed8", "none")))
}


custom_light <- function (...){
    data <- table_data()
    formatter("span",
              style = function(x) 
                style(display = "block", 
                      padding = "0 4px", 
                      `color` = "black", 
                      `border-radius` = "4px", 
                      `background-color` = ifelse(data$Incidence > 0.05, "#B1CBEB", "none")))
}


percent2 <- function(x, digits = 2, format = "f", ...) {
  paste0(formatC(1000 * x, format = format, digits = digits, ...), "%")
}

#bottom caption
# js <- c(
#   "function(settings){",
#   "  var datatable = settings.oInstance.api();",
#   "  var table = datatable.table().node();",
#   "  var caption = '*Rows highlighted in blue indicating the rates are extreme - more than 50 cases per 1000 persons-month.'",
#   "  $(table).append('<caption style=\"caption-side:bottom; text-align:left; font-size:11.5px;font-style:italic;color:#00CCFF;\">' + caption + '</caption>');",
#   "}"
# )

DT::renderDataTable({
  
  validate(
     need(nrow(table_data())>0, "Sorry, there is no data availale based on the indicated filter for this region.")
   )
  
  if(isTRUE(input$style)){
     data <- table_data() %>% formattable(list(area(col = 1:6) ~ custom_dark()))
  }else{
     data <- table_data() %>% formattable(list(area(col = 1:6) ~ custom_light()))
  }

  data$Incidence <- percent2(data$Incidence, 1)
  
  if(isTRUE(input$style)){
      my_table <- return(as.datatable(
     data, 
     fillContainer = TRUE,
     rownames = FALSE,
     colnames = c('Area', 'NZDep', 'Ethnicity', 'Population', 'Cases', 'Incidence rate*'),
     caption = tags$caption(
        tags$span("Table summarising filtered data where the incidence rate are larger than 5% (or 5 cases per 1000 persons-month) for", style = "color:white; margin:0 0 0 0; font-size:90%;"), tags$span(paste(input$Location, "."), style = 'color:#00CCFF;font-weight:bold; font-size:90%'), tags$br(),
        tags$span("*Rows highlighted in blue indicating the rates are extreme - more than 50 cases per 1000 persons-month.", style = "color:#2f7ed8;font-size:85%;font-style:italic;")
        ),
     options = list(
       #drawCallback = JS(js),
       searching = FALSE, paging = FALSE, lengthChange = FALSE, bInfo = FALSE,
       columnDefs = list(list(className = 'dt-left', targets = "_all")))
    ) %>%
      formatStyle('SA22018_Name', target = 'row', 
                  backgroundColor = "#34495e"))
  }else{
    my_table <- return(as.datatable(
     data, 
     fillContainer = TRUE,
     rownames = FALSE,
     colnames = c('Area', 'NZDep', 'Ethnicity', 'Population', 'Cases', 'Incidence rate*'),
     caption = tags$caption(
       tags$span("Table summarising filtered data where the incidence rate are larger than 5% (or 5 cases per 1000 persons-month) for", style = "color:black; margin:0 0 0 0; font-size:90%;"), tags$span(paste(input$Location, "."), style = 'color:#1a6ecc;font-weight:bold; font-size:90%'), tags$br(),
        tags$span("*Rows highlighted in blue indicating the rates are extreme - more than 50 cases per 1000 persons-month.", style = "color:#1a6ecc;font-size:85%;font-style:italic;")
        ),
     options = list(
       searching = FALSE, paging = FALSE, lengthChange = FALSE, bInfo = FALSE,
       columnDefs = list(list(className = 'dt-left', targets = "_all")))
    )
    )  
  }
  my_table
})
```
